---
title: "Preregistration - Main Analysis"
date: "`r format(Sys.time(), '%B %d, %Y')`"
author: "Isaac Chen"

format: 
    html:
      toc: true
      embed-resources: true
---

## Set up

```{r}
#| warning: false
#| message: false

library(here)
library(tidyverse)
library(interactions)
library(effects)
library(broom)

# Clear all variables from environment
rm(list = ls())

# Set input files
merg_dir  <- here("data", "phenotype", "merged")
participants_rds <- here(merg_dir, "preregistration-amendment_participants.rds")
  
# Read in data
participants_df <- readRDS(participants_rds)
```

## Staistical Models

### Three-way interactions

:::{.panel-tabset}

#### Model 1: TP ~ RS + FI + Age + (RS * FI) + (RS * Age) + (FI * Age) + (RS * FI * Age)

```{r}
# Model summary
lm1 <- lm(postC_TP ~ rs * fi * age, data = participants_df)

summary(lm1)
```

```{r}
#| echo: false
#| warning: false
#| message: false

# Interaction effect plot
lm1_effect <- interact_plot(
  model = lm1, 
  pred = rs, 
  modx = fi, 
  modx.values = c(70, 85, 100, 115, 130),
  mod2 = age, 
  mod2.values = c(10, 13.6),
  mod2.labels = c(
    "7.8 - 11.5 year olds", 
    "11.5 - 16.7 year olds"
  ),
  main.title = "Model 1: TP ~ RS * FI * Age",
  legend.main = "Fluid Intelligence",
  plot.points = T,
  colors = c("#d7191c", "#fdae61", "#fee08b", "#a6d96a", "#1a9641"),
  point.alpha = 0.8
)

fig_lm1_effect <- lm1_effect +
  scale_linetype_manual(
    name = "Fluid Intelligence",
    values = c("solid", "solid", "solid", "solid", "solid")
  ) + 
  scale_x_continuous(limits = c(55, 145), breaks = seq(55, 145, 10)) +
  scale_y_continuous(limits = c(0, 5), breaks = seq(0, 5, 1)) + 
  labs(
    x = "Reading Skill", 
    y = "Neural Basis of Phonological Processing (TP)",
  ) +
  theme_bw()

print(fig_lm1_effect)
```

```{r}
#| echo: false
#| eval: false
#| warning: false
#| message: false

# See below for how the data is split and plotted across panels
#'   Otherwise, a more ad hoc procedure is used to split the data. Quantiles
#'   are found for each `mod2.values` or `modx.values` value. These are not the
#'   quantiles used to split the data, however, since we want the plotted lines
#'   to represent the slope at a typical value in the group. The next step,
#'   then, is to take the mean of each pair of neighboring quantiles and use
#'   these as the cut points.
#'
#'   For example, if the `mod2.values` are at the 25th, 50th, and 75th
#'   percentiles
#'   of the distribution of the moderator, the data will be split at the
#'   37.5th and and 62.5th percentiles. When the variable is
#'   normally distributed, this will correspond fairly closely to using
#'   terciles.

mod2vals2 <- c(10, 13.6)
mod_val_qs <- ecdf(participants_df$age)(sort(mod2vals2))
cut_points2 <- c()
cut_points2 <- c()
for (i in 1:(length(mod2vals2) - 1)) {
  cut_points2 <- c(cut_points2, mean(mod_val_qs[i:(i + 1)]))
}
cut_points2 <- c(-Inf, quantile(participants_df$age, cut_points2, na.rm = TRUE), Inf)

gb <- ggplot_build(lm1_effect)
check <- gb$data[[2]] |> 
  rename(postC_TP = y)

check_df <- left_join(check, participants_df, by = "postC_TP") |> 
  relocate(age, .before = PANEL)
```

#### Model 2: vOT ~ PA + FI + Age + (PA * FI) + (PA * Age) + (FI * Age) + (PA * FI * Age)

```{r}
# Model summary
lm2 <- lm(postC_vOT ~ pa * fi * age, data = participants_df)

summary(lm2)
```

```{r}
#| echo: false
#| warning: false
#| message: false

# Interaction effect plot
lm2_effect <- interact_plot(
  model = lm2, 
  pred = pa, 
  modx = fi, 
  modx.values = c(70, 85, 100, 115, 130),
  mod2 = age, 
  mod2.values = c(10, 13.6),
  mod2.labels = c(
    "7.8 - 11.5 year olds", 
    "11.5 - 16.7 year olds"
  ),
  main.title = "Model 2: vOT ~ PA * FI * Age",
  legend.main = "Fluid Intelligence",
  plot.points = T,
  colors = c("#d7191c", "#fdae61", "#fee08b", "#a6d96a", "#1a9641"),
  point.alpha = 0.8
)

fig_lm2_effect <- lm2_effect +
  scale_linetype_manual(
    name = "Fluid Intelligence",
    values = c("solid", "solid", "solid", "solid", "solid")
  ) + 
  scale_x_continuous(limits = c(2, 19), breaks = seq(2, 19, 2)) + 
  scale_y_continuous(limits = c(0, 12), breaks = seq(0, 12, 2)) + 
  labs(
    x = "Phonological Awareness", 
    y = "Neural Basis of Orthographical Processing (vOT)",
  ) +
  theme_bw()

print(fig_lm2_effect)
```

#### Model 3: IFG ~ RS + FI + Age + (RS * FI) + (RS * Age) + (FI * Age) + (RS * FI * Age)

```{r}
# Model summary
lm3 <- lm(postC_IFG ~ rs * fi * age, data = participants_df)

summary(lm3)
```

```{r}
#| echo: false
#| warning: false
#| message: false

# Interaction effect plot
lm3_effect <- interact_plot(
  model = lm3, 
  pred = rs, 
  modx = fi, 
  modx.values = c(70, 85, 100, 115, 130),
  mod2 = age, 
  mod2.values = c(10, 13.6),
  mod2.labels = c(
    "7.8 - 11.5 year olds", 
    "11.5 - 16.7 year olds"
  ),
  main.title = "Model 3: IFG ~ RS * FI * Age",
  legend.main = "Fluid Intelligence",
  plot.points = T,
  colors = c("#d7191c", "#fdae61", "#fee08b", "#a6d96a", "#1a9641"),
  point.alpha = 0.8
)

fig_lm3_effect <- lm3_effect +
  scale_linetype_manual(
    name = "Fluid Intelligence",
    values = c("solid", "solid", "solid", "solid", "solid")
  ) + 
  scale_x_continuous(limits = c(55, 145), breaks = seq(55, 145, 10)) +
  scale_y_continuous(limits = c(0, 8.5), breaks = seq(0, 8, 2)) + 
  labs(
    x = "Reading Skill", 
    y = "Neural Basis of Lexical Processing (IFG)",
  ) +
  theme_bw()

print(fig_lm3_effect)
```

#### Model 4: IFG ~ PA + FI + Age + (PA * FI) + (PA * Age) + (FI * Age) + (PA * FI * Age)

```{r}
# Model summary
lm4 <- lm(postC_IFG ~ pa * fi * age, data = participants_df)

summary(lm4)
```

```{r}
#| echo: false
#| warning: false
#| message: false

# Interaction effect plot
lm4_effect <- interact_plot(
  model = lm4, 
  pred = pa, 
  modx = fi, 
  modx.values = c(70, 85, 100, 115, 130),
  mod2 = age, 
  mod2.values = c(10, 13.6),
  mod2.labels = c(
    "7.8 - 11.5 year olds", 
    "11.5 - 16.7 year olds"
  ),
  main.title = "Model 4: IFG ~ PA * FI * Age",
  legend.main = "Fluid Intelligence",
  plot.points = T,
  colors = c("#d7191c", "#fdae61", "#fee08b", "#a6d96a", "#1a9641"),
  point.alpha = 0.8
)

fig_lm4_effect <- lm4_effect +
  scale_linetype_manual(
    name = "Fluid Intelligence",
    values = c("solid", "solid", "solid", "solid", "solid")
  ) + 
  scale_x_continuous(limits = c(2, 19), breaks = seq(2, 19, 2)) + 
  scale_y_continuous(limits = c(0, 8.5), breaks = seq(0, 8, 2)) + 
  labs(
    x = "Phonological Awareness", 
    y = "Neural Basis of Lexical Processing (IFG)",
  ) +
  theme_bw()

print(fig_lm4_effect)
```

#### Model 5: vOT ~ RS + FI + Age + (RS * FI) + (RS * Age) + (FI * Age) + (RS * FI * Age)


```{r}
# Model summary
lm5 <- lm(postC_vOT ~ rs * fi * age, data = participants_df)

summary(lm5)
```

```{r}
#| echo: false
#| warning: false
#| message: false

# Interaction effect plot
lm5_effect <- interact_plot(
  model = lm5, 
  pred = rs, 
  modx = fi, 
  modx.values = c(70, 85, 100, 115, 130),
  mod2 = age, 
  mod2.values = c(10, 13.6),
  mod2.labels = c(
    "7.8 - 11.5 year olds", 
    "11.5 - 16.7 year olds"
  ),
  main.title = "Model 5: vOT ~ RS * FI * Age",
  legend.main = "Fluid Intelligence",
  plot.points = T,
  colors = c("#d7191c", "#fdae61", "#fee08b", "#a6d96a", "#1a9641"),
  point.alpha = 0.8
)

fig_lm5_effect <- lm5_effect +
  scale_linetype_manual(
    name = "Fluid Intelligence",
    values = c("solid", "solid", "solid", "solid", "solid")
  ) + 
  scale_x_continuous(limits = c(55, 145), breaks = seq(55, 145, 10)) +
  scale_y_continuous(limits = c(0, 12), breaks = seq(0, 12, 2)) + 
  labs(
    x = "Reading Skill", 
    y = "Neural Basis of Orthographic Processing (vOT)",
  ) +
  theme_bw()

print(fig_lm5_effect)
```

#### Model 6: TP ~ PA + FI + Age + (PA * FI) + (PA * Age) + (FI * Age) + (PA * FI * Age)

```{r}
# Model summary
lm6 <- lm(postC_TP ~ pa * fi * age, data = participants_df)

summary(lm6)
```

```{r}
#| echo: false
#| warning: false
#| message: false

# Interaction effect plot
lm6_effect <- interact_plot(
  model = lm6, 
  pred = pa, 
  modx = fi, 
  modx.values = c(70, 85, 100, 115, 130),
  mod2 = age, 
  mod2.values = c(10, 13.6),
  mod2.labels = c(
    "7.8 - 11.5 year olds", 
    "11.5 - 16.7 year olds"
  ),
  main.title = "Model 6: TP ~ PA * FI * Age",
  legend.main = "Fluid Intelligence",
  plot.points = T,
  colors = c("#d7191c", "#fdae61", "#fee08b", "#a6d96a", "#1a9641"),
  point.alpha = 0.8
)

fig_lm6_effect <- lm6_effect +
  scale_linetype_manual(
    name = "Fluid Intelligence",
    values = c("solid", "solid", "solid", "solid", "solid")
  ) + 
  scale_x_continuous(limits = c(2, 19), breaks = seq(2, 19, 2)) + 
  scale_y_continuous(limits = c(0, 5), breaks = seq(0, 5, 1)) + 
  labs(
    x = "Phonological Awareness", 
    y = "Neural Basis of Phonological Processing (TP)",
  ) +
  theme_bw()

print(fig_lm6_effect)
```

:::

### Two-way interactions

:::{.panel-tabset}

#### Model 7: TP ~ RS + FI + Age + (RS * FI)

```{r}
# Model summary
lm7 <- lm(postC_TP ~ rs * fi + age, data = participants_df)

summary(lm7)
```

```{r}
# Interaction effect plot
lm7_effect <- interact_plot(
  model = lm7, 
  pred = rs, 
  modx = fi, 
  modx.values = c(70, 85, 100, 115, 130),
  main.title = "Model 7: TP ~ RS + FI + Age + (RS * FI)",
  legend.main = "Fluid Intelligence",
  plot.points = T,
  colors = c("#d7191c", "#fdae61", "#fee08b", "#a6d96a", "#1a9641"),
  point.alpha = 0.8
)

fig_lm7_effect <- lm7_effect +
  scale_linetype_manual(
    name = "Fluid Intelligence",
    values = c("solid", "solid", "solid", "solid", "solid")
  ) + 
  scale_x_continuous(limits = c(55, 145), breaks = seq(55, 145, 10)) + 
  scale_y_continuous(limits = c(0, 5), breaks = seq(0, 5, 1)) + 
  labs(
    x = "Reading Skill", 
    y = "Neural Basis of Phonological Processing (TP)",
  ) + 
  theme_bw()

print(fig_lm7_effect)
```

#### Model 8: vOT ~ PA + FI + Age + (PA * FI)

```{r}
# Model summary
lm8 <- lm(postC_vOT ~ pa * fi + age, data = participants_df)

summary(lm8)
```

```{r}
# Interaction effect plot
lm8_effect <- interact_plot(
  model = lm8, 
  pred = pa, 
  modx = fi, 
  modx.values = c(70, 85, 100, 115, 130),
  main.title = "Model 8: vOT ~ PA + FI + Age + (PA * FI)",
  legend.main = "Fluid Intelligence",
  plot.points = T,
  colors = c("#d7191c", "#fdae61", "#fee08b", "#a6d96a", "#1a9641"),
  point.alpha = 0.8
)

fig_lm8_effect <- lm8_effect +
  scale_linetype_manual(
    name = "Fluid Intelligence",
    values = c("solid", "solid", "solid", "solid", "solid")
  ) + 
  scale_x_continuous(limits = c(2, 19), breaks = seq(2, 19, 2)) + 
  scale_y_continuous(limits = c(0, 12), breaks = seq(0, 12, 2)) +
  labs(
    x = "Phonological Awareness", 
    y = "Neural Basis of Orthographical Processing (vOT)",
  ) + 
  theme_bw()

print(fig_lm8_effect)
```

#### Model 9: IFG ~ RS + FI + Age + (RS * FI)

```{r}
# Model summary
lm9 <- lm(postC_IFG ~ rs * fi + age, data = participants_df)

summary(lm9)
```

```{r}
# Interaction effect plot
lm9_effect <- interact_plot(
  model = lm9, 
  pred = rs, 
  modx = fi, 
  modx.values = c(70, 85, 100, 115, 130),
  main.title = "Model 9: IFG ~ RS + FI + Age + (RS * FI)",
  legend.main = "Fluid Intelligence",
  plot.points = T,
  colors = c("#d7191c", "#fdae61", "#fee08b", "#a6d96a", "#1a9641"),
  point.alpha = 0.8
)

fig_lm9_effect <- lm9_effect +
  scale_linetype_manual(
    name = "Fluid Intelligence",
    values = c("solid", "solid", "solid", "solid", "solid")
  ) + 
  scale_x_continuous(limits = c(55, 145), breaks = seq(55, 145, 10)) +
  scale_y_continuous(limits = c(0, 8.5), breaks = seq(0, 8, 2)) +
  labs(
    x = "Reading Skill", 
    y = "Neural Basis of Lexical Processing (IFG)",
  ) + 
  theme_bw()

print(fig_lm9_effect)
```

#### Model 10: IFG ~ PA + FI + Age + (PA * FI)

```{r}
# Model summary
lm10 <- lm(postC_IFG ~ pa * fi + age, data = participants_df)

summary(lm10)
```

```{r}
# Interaction effect plot
lm10_effect <- interact_plot(
  model = lm10, 
  pred = pa, 
  modx = fi, 
  modx.values = c(70, 85, 100, 115, 130),
  main.title = "Model 10: IFG ~ PA + FI + Age + (PA * FI)",
  legend.main = "Fluid Intelligence",
  plot.points = T,
  colors = c("#d7191c", "#fdae61", "#fee08b", "#a6d96a", "#1a9641"),
  point.alpha = 0.8
)

fig_lm10_effect <- lm10_effect +
  scale_linetype_manual(
    name = "Fluid Intelligence",
    values = c("solid", "solid", "solid", "solid", "solid")
  ) + 
  scale_x_continuous(limits = c(2, 19), breaks = seq(2, 19, 2)) + 
  scale_y_continuous(limits = c(0, 8.5), breaks = seq(0, 8, 2)) + 
  labs(
    x = "Phonological Awareness", 
    y = "Neural Basis of Lexical Processing (IFG)",
  ) + 
  theme_bw()

print(fig_lm10_effect)
```

#### Model 11: vOT ~ RS + FI + Age + (RS * FI)

```{r}
# Model summary
lm11 <- lm(postC_vOT ~ rs * fi + age, data = participants_df)

summary(lm11)
```

```{r}
# Interaction effect plot
lm11_effect <- interact_plot(
  model = lm11, 
  pred = rs, 
  modx = fi, 
  modx.values = c(70, 85, 100, 115, 130),
  main.title = "Model 11: vOT ~ RS + FI + Age + (RS * FI)",
  legend.main = "Fluid Intelligence",
  plot.points = T,
  colors = c("#d7191c", "#fdae61", "#fee08b", "#a6d96a", "#1a9641"),
  point.alpha = 0.8
)

fig_lm11_effect <- lm11_effect +
  scale_linetype_manual(
    name = "Fluid Intelligence",
    values = c("solid", "solid", "solid", "solid", "solid")
  ) + 
  scale_x_continuous(limits = c(55, 145), breaks = seq(55, 145, 10)) +
  scale_y_continuous(limits = c(0, 12), breaks = seq(0, 12, 2)) +
  labs(
    x = "Reading Skill", 
    y = "Neural Basis of Orthographical Processing (vOT)",
  ) + 
  theme_bw()

print(fig_lm11_effect)
```

#### Model 12: TP ~ PA + FI + Age + (PA * FI)

```{r}
# Model summary
lm12 <- lm(postC_TP ~ pa * fi + age, data = participants_df)

summary(lm12)
```

```{r}
# Interaction effect plot
lm12_effect <- interact_plot(
  model = lm12, 
  pred = pa, 
  modx = fi, 
  modx.values = c(70, 85, 100, 115, 130),
  main.title = "Model 12: TP ~ PA + FI + Age + (PA * FI)",
  legend.main = "Fluid Intelligence",
  plot.points = T,
  colors = c("#d7191c", "#fdae61", "#fee08b", "#a6d96a", "#1a9641"),
  point.alpha = 0.8
)

fig_lm12_effect <- lm12_effect +
  scale_linetype_manual(
    name = "Fluid Intelligence",
    values = c("solid", "solid", "solid", "solid", "solid")
  ) + 
  scale_x_continuous(limits = c(2, 19), breaks = seq(2, 19, 2)) + 
  scale_y_continuous(limits = c(0, 5), breaks = seq(0, 5, 1)) + 
  labs(
    x = "Phonological Awareness", 
    y = "Neural Basis of Phonological Processing (TP)",
  ) + 
  theme_bw()

print(fig_lm12_effect)
```

:::

### Main effects

:::{.panel-tabset}

#### Model 13: TP ~ RS + FI + Age

```{r}
# Model summary
lm13 <- lm(postC_TP ~ rs + fi + age, data = participants_df)

summary(lm13)
```

```{r}
# Scatter plot
fig_lm13_scatter <- participants_df |> 
  ggplot(aes(x = rs, y = postC_TP)) +
  geom_point(aes(color = fi)) +
  geom_smooth(method = lm, formula = y ~ x, se = FALSE, color = "steelblue") +
  scale_color_gradientn(
    colors = c("#d7191c", "#fdae61", "#fee08b", "#a6d96a", "#1a9641")
  ) + 
  scale_x_continuous(limits = c(55, 145), breaks = seq(55, 145, 10)) + 
  scale_y_continuous(limits = c(0, 5), breaks = seq(0, 5, 1)) + 
  labs(
    title = "Model 13: TP ~ RS + FI + Age",
    x = "Reading Skill",
    y = "Neural Basis of Phonological Processing (TP)",
    color = "Fluid Intelligence"
  ) +
  theme_bw()

print(fig_lm13_scatter)
```

#### Model 14: vOT ~ PA + FI + Age

```{r}
# Model summary
lm14 <- lm(postC_vOT ~ pa + fi + age, data = participants_df)

summary(lm14)
```

```{r}
# Scatter plot
fig_lm14_scatter <- participants_df |> 
  ggplot(aes(x = pa, y = postC_vOT)) +
  geom_point(aes(color = fi)) +
  geom_smooth(method = lm, formula = y ~ x, se = FALSE, color = "steelblue") +
  scale_color_gradientn(
    colors = c("#d7191c", "#fdae61", "#fee08b", "#a6d96a", "#1a9641")
  ) +
  scale_x_continuous(limits = c(2, 19), breaks = seq(2, 19, 2)) + 
  scale_y_continuous(limits = c(0, 12), breaks = seq(0, 12, 2)) +
  labs(
    title = "Model 14: vOT ~ PA + FI + Age",
    x = "Phonological Awareness",
    y = "Neural Basis of Orthographical Processing (vOT)",
    color = "Fluid Intelligence"
  ) +
  theme_bw()

print(fig_lm14_scatter)
```

#### Model 15: IFG ~ RS + FI + Age

```{r}
# Model summary
lm15 <- lm(postC_IFG ~ rs + fi + age, data = participants_df)

summary(lm15)
```

```{r}
# Scatter plot
fig_lm15_scatter <- participants_df |> 
  ggplot(aes(x = rs, y = postC_IFG)) +
  geom_point(aes(color = fi)) +
  geom_smooth(method = lm, formula = y ~ x, se = FALSE, color = "steelblue") +
  scale_color_gradientn(
    colors = c("#d7191c", "#fdae61", "#fee08b", "#a6d96a", "#1a9641")
  ) +
  scale_x_continuous(limits = c(55, 145), breaks = seq(55, 145, 10)) +
  scale_y_continuous(limits = c(0, 8.5), breaks = seq(0, 8, 2)) +
  labs(
    title = "Model 15: IFG ~ RS + FI + Age",
    x = "Reading Skill",
    y = "Neural Basis of Lexical Processing (IFG)",
    color = "Fluid Intelligence"
  ) +
  theme_bw()

print(fig_lm15_scatter)
```

#### Model 16: IFG ~ PA + FI + Age

```{r}
# Model summary
lm16 <- lm(postC_IFG ~ pa + fi + age, data = participants_df)

summary(lm16)
```

```{r}
# Scatter plot
fig_lm16_scatter <- participants_df |> 
  ggplot(aes(x = pa, y = postC_IFG)) +
  geom_point(aes(color = fi)) +
  geom_smooth(method = lm, formula = y ~ x, se = FALSE, color = "steelblue") +
  scale_color_gradientn(
    colors = c("#d7191c", "#fdae61", "#fee08b", "#a6d96a", "#1a9641")
  ) +
  scale_x_continuous(limits = c(2, 19), breaks = seq(2, 19, 2)) + 
  scale_y_continuous(limits = c(0, 8.5), breaks = seq(0, 8, 2)) +
  labs(
    title = "Model 16: IFG ~ PA + FI + Age",
    x = "Phonological Awareness",
    y = "Neural Basis of Lexical Processing (IFG)",
    color = "Fluid Intelligence"
  ) +
  theme_bw()

print(fig_lm16_scatter)
```

#### Model 17: vOT ~ RS + FI + Age

```{r}
# Model summary
lm17 <- lm(postC_vOT ~ rs + fi + age, data = participants_df)

summary(lm17)
```

```{r}
# Scatter plot
fig_lm17_scatter <- participants_df |> 
  ggplot(aes(x = rs, y = postC_vOT)) +
  geom_point(aes(color = fi)) +
  geom_smooth(method = lm, formula = y ~ x, se = FALSE, color = "steelblue") +
  scale_color_gradientn(
    colors = c("#d7191c", "#fdae61", "#fee08b", "#a6d96a", "#1a9641")
  ) +
  scale_x_continuous(limits = c(55, 145), breaks = seq(55, 145, 10)) +
  scale_y_continuous(limits = c(0, 12), breaks = seq(0, 12, 2)) +
  labs(
    title = "Model 17: vOT ~ RS + FI + Age",
    x = "Reading Skill",
    y = "Neural Basis of Orthographical Processing (vOT)",
    color = "Fluid Intelligence"
  ) +
  theme_bw()

print(fig_lm17_scatter)
```

#### Model 18: TP ~ PA + FI + Age

```{r}
# Model summary
lm18 <- lm(postC_TP ~ pa + fi + age, data = participants_df)

summary(lm18)
```

```{r}
# Scatter plot
fig_lm18_scatter <- participants_df |> 
  ggplot(aes(x = pa, y = postC_TP)) +
  geom_point(aes(color = fi)) +
  geom_smooth(method = lm, formula = y ~ x, se = FALSE, color = "steelblue") +
  scale_color_gradientn(
    colors = c("#d7191c", "#fdae61", "#fee08b", "#a6d96a", "#1a9641")
  ) + 
  scale_x_continuous(limits = c(2, 19), breaks = seq(2, 19, 2)) + 
  scale_y_continuous(limits = c(0, 5), breaks = seq(0, 5, 1)) + 
  labs(
    title = "Model 18: TP ~ PA + FI + Age",
    x = "Phonological Awareness",
    y = "Neural Basis of Phonological Processing (TP)",
    color = "Fluid Intelligence"
  ) +
  theme_bw()

print(fig_lm18_scatter)
```

#### FDR correction

```{r}
lm13_pval <- tidy(lm13) |> filter(term == "rs") |> pull(p.value)
lm14_pval <- tidy(lm14) |> filter(term == "pa") |> pull(p.value)
lm15_pval <- tidy(lm15) |> filter(term == "rs") |> pull(p.value)
lm16_pval <- tidy(lm16) |> filter(term == "pa") |> pull(p.value)
lm17_pval <- tidy(lm17) |> filter(term == "rs") |> pull(p.value)
lm18_pval <- tidy(lm18) |> filter(term == "pa") |> pull(p.value)

pvals <- c(lm13_pval, lm14_pval, lm15_pval, lm16_pval, lm17_pval, lm18_pval)

pvals_fdr <- p.adjust(pvals, method = "fdr")
print(pvals_fdr)
```

:::
