---
title: "Create ds001486 Dataset"
date: "`r format(Sys.time(), '%B %d, %Y')`"
author: "Isaac Chen"

format: 
    html:
      embed-resources: true
      code-link: true
---

### About
This dataset was created using `ds001486` `version 1.3.1` ([Link to OpenNeuro](https://openneuro.org/datasets/ds001486/versions/1.3.1)).

This script aims to create the following two files:

- `{dataset}_participants.csv`
  - Provides information such as age, sex, handedness, and assessment outcomes
- `{dataset}_task-WordRhyming_events.csv`
  - Provides information about item-level performance on in-scanner task across
  all participant scans

### Known Issues

- 133 bold scans contain a different number of volumes due to an adjustment in the number occuring after the task had completed
- Inproper referencing of the correct answer caused an error in the accuracy and response time reporting for some trials. These trials are marked as "err" for accuracy and response_time in the events file
- task-Num_run-02 and task-Sub_run-01 at both ses-T1 and ses-T2 only allowed participants to respond for 800 ms following presentation of the control, blue box, stimulus. This issue only applied to the control trial_type, and may have resulted in lower accuracy and higher non-response for this condition, in these runs.

### Note

- To address trial/rows marked as "err" in the events file, the `condition` column will keep track of this information to allow the `accuracy` and `response_time` columns to remain as numeric data types.
- These "err" trials will be removed first when calculating accuracy and response bias later on for those participants.

### Change Log:
- 2024-09-13: Isaac updated file path to work on his local computer
- 2024-10-07: Isaac updated to check if data matches `Suárez-Pellicioni et al. 2019`
- 2025-10-10: Isaac added documentation and streamlined script

### Set up

```{r}
#| warning: false
#| message: false
# Clear all variables from environment
rm(list = ls())

# Load library
library(here)
library(tidyverse)
library(glue)

# Name dataset of interest
dataset = "ds001486"

# Set the directory and files
raw_dir  <- here("data", "phenotype", "raw", dataset)
proc_dir <- here("data", "phenotype", "processed", dataset)
phen_dir <- here(raw_dir, "phenotype", "ses-T1")

# Set input files
participants_tsv <- here(raw_dir , "participants.tsv") 
dev_history_tsv  <- here(phen_dir, "developmental_history_questionnaire.tsv")
wj3_tsv          <- here(phen_dir, "wj-III.tsv")
ctopp_tsv        <- here(phen_dir, "ctopp.tsv")
wasi_tsv         <- here(phen_dir, "wasi.tsv")

# Set output files
output_participants   <- here(proc_dir, glue("{dataset}_participants.csv"))
output_rhyming_events <- here(proc_dir, glue("{dataset}_task-VisualWordRhyming_events.csv"))
```

### Process participant summary and assessment data

```{r}
#| warning: false
#| message: false
# Read in participant summary data
participants_df <- read_tsv(participants_tsv)
dev_history_df  <- read_tsv(dev_history_tsv)

# Select variables from developmental history survey
dev_history_df <- dev_history_df |> 
  select(participant_id, lives_with, ends_with("_highest_grade"))

# Combine two datasets into one
# Rename sex and code handedness
participants_df <- participants_df |> 
  select(participant_id, `age_ses-T1`, sex, handedness) |> 
  right_join(dev_history_df, by = "participant_id") |> 
  mutate(
    sex = case_when(
      sex == 1 ~ "male",
      sex == 2 ~ "female",
      .default = NA_character_
    ),
    handedness = case_when(
      handedness == 1 ~ "left",
      handedness == 2 ~ "right",
      .default = NA_character_
    )
  )
```

```{r}
#| warning: false
#| message: false
# Read in assessment data
wj3_df   <- read_tsv(wj3_tsv) 
ctopp_df <- read_tsv(ctopp_tsv) 
wasi_df  <- read_tsv(wasi_tsv) 

# Append assessment data
participants_df <- participants_df |> 
  right_join(wj3_df,   by = "participant_id") |> 
  right_join(ctopp_df, by = "participant_id") |> 
  right_join(wasi_df,  by = "participant_id")
```

```{r}
# Process record IDs to include unique ID and dataset
participants_df <- participants_df |>
  mutate(
    id        = as.numeric(stringr::str_remove_all(participant_id, "sub-")),
    unique_id = glue("{dataset}_sub-{formatC(id, width = 4, flag = 0)}"),
    dataset   = dataset,
    .before   = participant_id
  ) |> 
  select(-id)
```

### Search for word rhyming task events file

```{r}
# Define the regular expression pattern for word rhyming task
pattern <- ".*ses-T1_task-Rhyming.*tsv"

# Use list.files with pattern argument to list files matching the regex
rhyming_events_files <- 
  list.files(
    path = raw_dir, 
    pattern = pattern, 
    full.names = TRUE, 
    recursive = TRUE
  )

# List number of files found
glue("Found {length(rhyming_events_files)} word rhyming task events files")
```

### Description of trial

| Name         | Number | Description                                                   | 
|--------------|--------|---------------------------------------------------------------|
|1. O+P+       | 12     | Matching words, shared phonology, shared orthography          |
|2. O+P-       | 10     | Matching words, shared phonology, no shared orthography       |
|3. O-P+       | 10     | Mismatching words, no shared phonology, shared orthography    |
|4. O-P-       | 14     | Mismatching words, no shared phonology, no shared orthography |
|5. Control    | 25     | Respond with index finger when blue square turns red          |
|6. Perceptual | 13     | Matching judgment of non-alphabetic characters                |
|NA. Red box   | 1      | Red square presented at the end of stimulus presentation      |
|**Total:**    | **85** |                                                               |
: {tbl-colwidths="[20, 10, 70]" .striped}

Source: `task-Rhyming_bold.json`

- !!!The red box trial will be removed. So total trial number is 84!!!

### Function to extract data from events file

This function will create a events database with the following columns:

1. `unique_id`: 
2. `dataset`
3. `participant_id`
4. `session`
5. `task`
6. `acquisition`
7. `run`
8. `onset`
9. `duration`
10. `trial_type`
11. `trial_name`
12. `condition`
13. `stim1`
14. `stim2`
15. `response`
16. `accuracy`

```{r}
# Define function for extracting events data
extract_events_data <- function(events_file, dataset, events_num) {
  # For troubleshooting purposes
  #events_file = rhyming_events_files[1]
  
  # Extract file name and break it up into parts
  file_name       <- str_remove_all(events_file, "^.*/")
  file_name_parts <- str_split_1(file_name, "_")
  
  # Extract subject ID, session, run
  current_subject     <- file_name_parts[1]
  current_id          <- file_name_parts[1] |> stringr::str_remove_all("sub-")
  current_session     <- file_name_parts[2]
  current_task        <- file_name_parts[3]
  #current_acquisition <- file_name_parts[4]
  #current_run         <- file_name_parts[5]    #ds001486 only contains 1 run
  
  # Read in events TSV file
  # Make sure datatype is consistent
  # Remove Red box row
  events <- 
    read_tsv(events_file, show_col_types = FALSE) |> 
    rename(
      stim1 = prime_stim,
      stim2 = target_stim
    ) |> 
    mutate(
      onset         = as.character(onset),      # as.character to avoid warnings
      duration      = as.character(duration),
      trial_type    = as.character(trial_type),
      #trial_name    = as.character(trial_name) ,
      stim1         = as.character(stim1),
      stim2         = as.character(stim2),
      response_time = as.character(response_time),
      accuracy      = as.character(accuracy)
    ) |> 
    filter(stim2 != "/REDbox.bmp")
    
  
  # Check number of rows in events file
  if (nrow(events) != events_num) {
    stop(glue("Error: expected {events_num} rows, but found {nrow(events)}"))
  }
  
  # Add metadata to events file information
  subject_data <- events |> 
    mutate(
      # Metadata
      unique_id      = glue("{dataset}_sub-{formatC(as.numeric(current_id), width = 4, flag = 0)}"),
      dataset        = dataset,
      participant_id = current_subject,
      session        = current_session,
      task           = current_task,
      acquisition    = "",
      run            = "run-01",
      
      # Categorize trials 
      condition = case_when(
        trial_type == "1.0"  ~ "O+P+",
        trial_type == "2.0"  ~ "O+P-",       
        trial_type == "3.0"  ~ "O-P+",
        trial_type == "4.0"  ~ "O-P-",
        trial_type == "5.0"  ~ "control",
        trial_type == "6.0"  ~ "perceptual",
      ),
      
      # Address "err" trials
      condition = ifelse(accuracy == "err" | response_time == "err", "err", condition)
      
    )
  
  return(subject_data)
}
```

### Create word rhyming task events dataset

```{r}
#| eval: false
# Previous method wtih purr::map_dfr
#rhyming.df <- map_dfr(.x = rhyming_events_files, .f = ~ clean_subj_data(.x))

# Print warnings immediately
options(warn = 1)
#options(warn = 0) # return to default

# Set parameters  
rhyming_events_df <- tibble()
dataset = "ds001486"
events_num = 84

# Create word rhyming in-scanner task dataset
for (idx in seq_along(rhyming_events_files)) {
  # Extract file name
  file_name <- str_remove_all(rhyming_events_files[idx], "^.*/")
  cat(glue("{idx}. Cleaning {file_name}..."))
  cat(glue("done"), sep = "\n")
  
  # Extract single scan events data
  subject_events <- extract_events_data(
    events_file = rhyming_events_files[idx], 
    dataset = dataset,
    events_num = events_num
  )
  
  # Append single scan events data to dataset
  rhyming_events_df <- bind_rows(rhyming_events_df, subject_events)
}

# Reorder columns for clarity
# Convert columns to numeric
# If resp is missing (NA), then make acc = 0 (incorrect)
rhyming_events_df <- rhyming_events_df |>
  select(
    unique_id:run, onset:trial_type, condition, response_time, accuracy,
    stim1, stim2
  ) |> 
  mutate(
    onset         = as.numeric(onset),
    duration      = round(as.numeric(duration), 3),
    trial_type    = as.numeric(trial_type),
    response_time = as.numeric(response_time),
    accuracy      = as.numeric(accuracy)
  )
```

### Quality Check

```{r}
#| eval: false
# Inspect any rows that have NAs
qc <- rhyming_events_df |>filter(is.na(accuracy))
glimpse(qc)

# Change accuracy to 0 (incorrect)
rhyming_events_df <- rhyming_events_df |> 
  mutate(accuracy = ifelse(is.na(accuracy), 0, accuracy))

# Count number of NAs across all columns
qc <- rhyming_events_df |>filter(if_any(everything(), is.na))
glimpse(qc)

# Count number of "err" trials
err_df <- rhyming_events_df |> filter(condition == "err")
dim(err_df)
```

### Save outputs

```{r}
#| eval: false
# Write outputs to csv files
write_csv(participants_df,   file = output_participants)
write_csv(rhyming_events_df, file = output_rhyming_events)
```
