---
title: "Preregistration - Exploratory Analysis"
date: "`r format(Sys.time(), '%B %d, %Y')`"
author: "Isaac Chen"

format: 
    html:
      toc: true
      embed-resources: true
---

## Set up

```{r}
#| warning: false
#| message: false

library(here)
library(tidyverse)
library(interactions)
library(effects)

# Clear all variables from environment
rm(list = ls())

# Set input files
merg_dir  <- here("data", "phenotype", "merged")
participants_rds <- here(merg_dir, "preregistration-amendment_participants.rds")
  
# Read in data
participants_df <- readRDS(participants_rds)
```

## Staistical Models

### Three-way interactions

:::{.panel-tabset}

#### Model 1: RS ~ PA + FI + Age + Dataset + (PA * FI) + (PA * Age) + (FI * Age) + (PA * FI * Age)

```{r}
# Model summary
lm1 <- lm(rs ~ pa * fi * age + dataset, data = participants_df)

summary(lm1)
```

```{r}
#| echo: false
#| warning: false
#| message: false

# Interaction effect plot
lm1_effect <- interact_plot(
  model = lm1, 
  pred = pa, 
  modx = fi, 
  modx.values = c(70, 85, 100, 115, 130),
  mod2 = age, 
  mod2.values = c(10, 13.6),
  mod2.labels = c(
    "7.8 - 11.5 year olds", 
    "11.5 - 16.7 year olds"
  ),
  main.title = "Model 1: RS ~ PA * FI * Age + Dataset",
  legend.main = "Fluid Intelligence",
  plot.points = T,
  colors = c("#d7191c", "#fdae61", "#fee08b", "#a6d96a", "#1a9641"),
  point.alpha = 0.8
)

fig_lm1_effect <- lm1_effect +
  scale_linetype_manual(
    name = "Fluid Intelligence",
    values = c("solid", "solid", "solid", "solid", "solid")
  ) + 
  scale_x_continuous(limits = c(2, 19), breaks = seq(2, 19, 2)) + 
  scale_y_continuous(limits = c(55, 145), breaks = seq(55, 145, 10)) +
  labs(
    x = "Phonological Awareness", 
    y = "Reading Skill",
  ) +
  theme_bw()

print(fig_lm1_effect)
```

#### Model 2: PA ~ RS + FI + Age + Dataset + (RS * FI) + (RS * Age) + (FI * Age) + (RS * FI * Age)

```{r}
# Model summary
lm2 <- lm(pa ~ rs * fi * age + dataset, data = participants_df)

summary(lm2)
```

```{r}
#| echo: false
#| warning: false
#| message: false

# Interaction effect plot
lm2_effect <- interact_plot(
  model = lm2, 
  pred = rs, 
  modx = fi, 
  modx.values = c(70, 85, 100, 115, 130),
  mod2 = age, 
  mod2.values = c(10, 13.6),
  mod2.labels = c(
    "7.8 - 11.5 year olds", 
    "11.5 - 16.7 year olds"
  ),
  main.title = "Model 2: PA ~ RS * FI * Age + Dataset",
  legend.main = "Fluid Intelligence",
  plot.points = T,
  colors = c("#d7191c", "#fdae61", "#fee08b", "#a6d96a", "#1a9641"),
  point.alpha = 0.8
)

fig_lm2_effect <- lm2_effect +
  scale_linetype_manual(
    name = "Fluid Intelligence",
    values = c("solid", "solid", "solid", "solid", "solid")
  ) +
  scale_x_continuous(limits = c(55, 145), breaks = seq(55, 145, 10)) +
  scale_y_continuous(limits = c(2, 19), breaks = seq(2, 19, 2)) + 
  labs(
    x = "Reading Skill", 
    y = "Phonological Awareness",
  ) +
  theme_bw()

print(fig_lm2_effect)
```

#### Model 3: ACC ~ PA + FI + Age + Dataset + (PA * FI) + (PA * Age) + (FI * Age) + (PA * FI * Age)

```{r}
# Model summary
lm3 <- lm(nci_acc ~ pa * fi * age + dataset, data = participants_df)

summary(lm3)
```

```{r}
#| echo: false
#| warning: false
#| message: false

# Interaction effect plot
lm3_effect <- interact_plot(
  model = lm3, 
  pred = pa, 
  modx = fi, 
  modx.values = c(70, 85, 100, 115, 130),
  mod2 = age, 
  mod2.values = c(10, 13.6),
  mod2.labels = c(
    "7.8 - 11.5 year olds", 
    "11.5 - 16.7 year olds"
  ),
  main.title = "Model 3: ACC ~ PA * FI * Age + Dataset",
  legend.main = "Fluid Intelligence",
  plot.points = T,
  colors = c("#d7191c", "#fdae61", "#fee08b", "#a6d96a", "#1a9641"),
  point.alpha = 0.8
)

fig_lm3_effect <- lm3_effect +
  scale_linetype_manual(
    name = "Fluid Intelligence",
    values = c("solid", "solid", "solid", "solid", "solid")
  ) +
  scale_x_continuous(limits = c(2, 19), breaks = seq(2, 19, 2)) + 
  scale_y_continuous(limits = c(0.5, 1), breaks = seq(0.5, 1, 0.1)) + 
  labs(
    x = "Phonological Awareness", 
    y = "Accuracy on NCIs",
  ) +
  theme_bw()

print(fig_lm3_effect)
```

#### Model 4: ACC ~ RS + FI + Age + Dataset + (RS * FI) + (RS * Age) + (FI * Age) + (RS * FI * Age)

```{r}
# Model summary
lm4 <- lm(nci_acc ~ rs * fi * age + dataset, data = participants_df)

summary(lm4)
```

```{r}
#| echo: false
#| warning: false
#| message: false

# Interaction effect plot
lm4_effect <- interact_plot(
  model = lm4, 
  pred = rs, 
  modx = fi, 
  modx.values = c(70, 85, 100, 115, 130),
  mod2 = age, 
  mod2.values = c(10, 13.6),
  mod2.labels = c(
    "7.8 - 11.5 year olds", 
    "11.5 - 16.7 year olds"
  ),
  main.title = "Model 4: ACC ~ RS * FI * Age + Dataset",
  legend.main = "Fluid Intelligence",
  plot.points = T,
  colors = c("#d7191c", "#fdae61", "#fee08b", "#a6d96a", "#1a9641"),
  point.alpha = 0.8
)

fig_lm4_effect <- lm4_effect +
  scale_linetype_manual(
    name = "Fluid Intelligence",
    values = c("solid", "solid", "solid", "solid", "solid")
  ) +
  scale_x_continuous(limits = c(55, 145), breaks = seq(55, 145, 10)) +
  scale_y_continuous(limits = c(0.5, 1), breaks = seq(0.5, 1, 0.1)) + 
  labs(
    x = "Reading Skill", 
    y = "Accuracy on NCIs",
  ) +
  theme_bw()

print(fig_lm4_effect)
```

:::

### Two-way interactions

:::{.panel-tabset}

#### Model 5: RS ~ PA + FI + Age + Dataset + (PA * FI)

```{r}
# Model summary
lm5 <- lm(rs ~ pa * fi + age + dataset, data = participants_df)

summary(lm5)
```

```{r}
# Interaction effect plot
lm5_effect <- interact_plot(
  model = lm5, 
  pred = pa, 
  modx = fi, 
  modx.values = c(70, 85, 100, 115, 130),
  main.title = "Model 5: RS ~ PA + FI + Age + Dataset + (PA * FI)",
  legend.main = "Fluid Intelligence",
  plot.points = T,
  colors = c("#d7191c", "#fdae61", "#fee08b", "#a6d96a", "#1a9641"),
  point.alpha = 0.8
)

fig_lm5_effect <- lm5_effect +
  scale_linetype_manual(
    name = "Fluid Intelligence",
    values = c("solid", "solid", "solid", "solid", "solid")
  ) + 
  scale_x_continuous(limits = c(2, 19), breaks = seq(2, 19, 2)) + 
  scale_y_continuous(limits = c(55, 145), breaks = seq(55, 145, 10)) +
  labs(
    x = "Phonological Awareness", 
    y = "Reading Skill",
  ) + 
  theme_bw()

print(fig_lm5_effect)
```

#### Model 6: PA ~ RS + FI + Age + Dataset + (RS * FI)

```{r}
# Model summary
lm6 <- lm(pa ~ rs * fi + age + dataset, data = participants_df)

summary(lm6)
```

```{r}
# Interaction effect plot
lm6_effect <- interact_plot(
  model = lm6, 
  pred = rs, 
  modx = fi, 
  modx.values = c(70, 85, 100, 115, 130),
  main.title = "Model 6: PA ~ RS + FI + Age + Dataset + (RS * FI)",
  legend.main = "Fluid Intelligence",
  plot.points = T,
  colors = c("#d7191c", "#fdae61", "#fee08b", "#a6d96a", "#1a9641"),
  point.alpha = 0.8
)

fig_lm6_effect <- lm6_effect +
  scale_linetype_manual(
    name = "Fluid Intelligence",
    values = c("solid", "solid", "solid", "solid", "solid")
  ) + 
  scale_x_continuous(limits = c(55, 145), breaks = seq(55, 145, 10)) +
  scale_y_continuous(limits = c(2, 19), breaks = seq(2, 19, 2)) + 
  labs(
    x = "Reading Skill", 
    y = "Phonological Awareness",
  ) + 
  theme_bw()

print(fig_lm6_effect)
```

#### Model 7: ACC ~ PA + FI + Age + Dataset + (PA * FI)

```{r}
# Model summary
lm7 <- lm(nci_acc ~ pa * fi + age + dataset, data = participants_df)

summary(lm7)
```

```{r}
#| echo: false
#| warning: false
#| message: false

# Interaction effect plot
lm7_effect <- interact_plot(
  model = lm7, 
  pred = pa, 
  modx = fi, 
  modx.values = c(70, 85, 100, 115, 130),
  main.title = "Model 7: ACC ~ PA + FI + Age + Dataset + (PA * FI)",
  legend.main = "Fluid Intelligence",
  plot.points = T,
  colors = c("#d7191c", "#fdae61", "#fee08b", "#a6d96a", "#1a9641"),
  point.alpha = 0.8
)

fig_lm7_effect <- lm7_effect +
  scale_linetype_manual(
    name = "Fluid Intelligence",
    values = c("solid", "solid", "solid", "solid", "solid")
  ) +
  scale_x_continuous(limits = c(2, 19), breaks = seq(2, 19, 2)) + 
  scale_y_continuous(limits = c(0.5, 1.0), breaks = seq(0.5, 1, 0.1)) + 
  labs(
    x = "Phonological Awareness", 
    y = "Accuracy on NCIs",
  ) +
  theme_bw()

print(fig_lm7_effect)
```

#### Model 8: ACC ~ RS + FI + Age + Dataset + (RS * FI)

```{r}
# Model summary
lm8 <- lm(nci_acc ~ rs * fi + age + dataset, data = participants_df)

summary(lm8)
```

```{r}
#| echo: false
#| warning: false
#| message: false

# Interaction effect plot
lm8_effect <- interact_plot(
  model = lm8, 
  pred = rs, 
  modx = fi, 
  modx.values = c(70, 85, 100, 115, 130),
  main.title = "Model 8: ACC ~ RS + FI + Age + Dataset + (RS * FI)",
  legend.main = "Fluid Intelligence",
  plot.points = T,
  colors = c("#d7191c", "#fdae61", "#fee08b", "#a6d96a", "#1a9641"),
  point.alpha = 0.8
)

fig_lm8_effect <- lm8_effect +
  scale_linetype_manual(
    name = "Fluid Intelligence",
    values = c("solid", "solid", "solid", "solid", "solid")
  ) +
  scale_x_continuous(limits = c(55, 145), breaks = seq(55, 145, 10)) +
  scale_y_continuous(limits = c(0.5, 1), breaks = seq(0.5, 1, 0.1)) + 
  labs(
    x = "Reading Skill", 
    y = "Accuracy on NCIs",
  ) +
  theme_bw()

print(fig_lm8_effect)
```

:::

### Main Effects

:::{.panel-tabset}

#### Model 9: RS ~ PA + FI + Age + Dataset

```{r}
# Model summary
lm9 <- lm(rs ~ pa + fi + age + dataset, data = participants_df)

summary(lm9)
```

```{r}
# Scatter plot
fig_lm9_scatter <- participants_df |> 
  ggplot(aes(x = pa, y = rs)) +
  geom_point(aes(color = fi)) +
  geom_smooth(method = lm, formula = y ~ x, se = FALSE, color = "steelblue") +
  scale_color_gradientn(
    colors = c("#d7191c", "#fdae61", "#fee08b", "#a6d96a", "#1a9641")
  ) +
  scale_x_continuous(limits = c(2, 19), breaks = seq(2, 19, 2)) + 
  scale_y_continuous(limits = c(55, 145), breaks = seq(55, 145, 10)) +
  labs(
    title = "Model 9: RS ~ PA + FI + Age + Dataset",
    x = "Phonological Awareness",
    y = "Reading Skill",
    color = "Fluid Intelligence"
  ) +
  theme_bw()

print(fig_lm9_scatter)
```

#### Model 10: PA ~ RS + FI + Age + Dataset

```{r}
# Model summary
lm10 <- lm(pa ~ rs + fi + age + dataset, data = participants_df)

summary(lm10)
```

```{r}
# Scatter plot
fig_lm10_scatter <- participants_df |> 
  ggplot(aes(x = rs, y = pa)) +
  geom_point(aes(color = fi)) +
  geom_smooth(method = lm, formula = y ~ x, se = FALSE, color = "steelblue") +
  scale_color_gradientn(
    colors = c("#d7191c", "#fdae61", "#fee08b", "#a6d96a", "#1a9641")
  ) + 
  scale_x_continuous(limits = c(55, 145), breaks = seq(55, 145, 10)) +
  scale_y_continuous(limits = c(2, 19), breaks = seq(2, 19, 2)) + 
  labs(
    title = "Model 10: PA ~ RS + FI + Age + Dataset",
    x = "Reading Skill",
    y = "Phonological Awareness",
    color = "Fluid Intelligence"
  ) +
  theme_bw()

print(fig_lm10_scatter)
```

#### Model 11: ACC ~ PA + FI + Age + Dataset

```{r}
# Model summary
lm11 <- lm(nci_acc ~ pa + fi + age + dataset, data = participants_df)

summary(lm11)
```

```{r}
# Scatter plot
fig_lm11_scatter <- participants_df |> 
  ggplot(aes(x = pa, y = nci_acc)) +
  geom_point(aes(color = fi)) +
  geom_smooth(method = lm, formula = y ~ x, se = FALSE, color = "steelblue") +
  scale_color_gradientn(
    colors = c("#d7191c", "#fdae61", "#fee08b", "#a6d96a", "#1a9641")
  ) +
  scale_x_continuous(limits = c(2, 19), breaks = seq(2, 19, 2)) + 
  scale_y_continuous(limits = c(0.5, 1.0), breaks = seq(0.5, 1, 0.1)) + 
  labs(
    title = "Model 11: ACC ~ PA + FI + Age + Dataset",
    x = "Phonological Awareness",
    y = "Accuracy on NCIs",
    color = "Fluid Intelligence"
  ) +
  theme_bw()

print(fig_lm11_scatter)
```

#### Model 12: ACC ~ RS + FI + Age + Dataset

```{r}
# Model summary
lm12 <- lm(nci_acc ~ rs + fi + age + dataset, data = participants_df)

summary(lm12)
```

```{r}
# Scatter plot
fig_lm12_scatter <- participants_df |> 
  ggplot(aes(x = rs, y = nci_acc)) +
  geom_point(aes(color = fi)) +
  geom_smooth(method = lm, formula = y ~ x, se = FALSE, color = "steelblue") +
  scale_color_gradientn(
    colors = c("#d7191c", "#fdae61", "#fee08b", "#a6d96a", "#1a9641")
  ) +
  scale_x_continuous(limits = c(55, 145), breaks = seq(55, 145, 10)) +
  scale_y_continuous(limits = c(0.5, 1), breaks = seq(0.5, 1, 0.1)) + 
  labs(
    title = "Model 12: ACC ~ RS + FI + Age + Dataset",
    x = "Reading Skill",
    y = "Accuracy on NCIs",
    color = "Fluid Intelligence"
  ) +
  theme_bw()

print(fig_lm12_scatter)
```

:::
